---
title: "101AHW6"
author: "JING LI"
date: "11/13/2016"
output: pdf_document
---

Statistics 101A- Fall 2016 Homework six
Part one: Using the UCLA stress echo data set (this data set and its description is posted on week eight of CCLE).
Using R conduct two-way ANOVA using maximum heart rate as the outcome variable, gender with diagnosis as the independent variables. Include your output.

```{r}
stress <- read.delim("~/Downloads/Stress2.txt", stringsAsFactors=FALSE) 
attach(stress)
boxplot(maxhr ~ gender, main="by Gender")
lm <- lm(maxhr~gender+newPTCA+gender:newPTCA)
anova(lm)
```


a) State the null and the alternative hypotheses within context.

H0: The means for female and male maximum heart rate is the same. The means for patients who had a diagnosis is the same.
Ha: The means for female and male maximum heart rate is not the same. The means for patients who had a diagnosis is not the same.

b) Based on the R output decide which hypotheses you reject and which ones
you fail to reject. Interpret the results within context.

Base on the p-value, none of the predictors are significant. We fail to reject null. There is no significant difference between the means for predictors gender and newPTCA.

c) Using R, create table of means and the plot of interaction effect between
gender and diagnosis. Interpret your conclusions regarding the interaction
effect within context.
```{r}
lmc <- aov(maxhr~gender+newPTCA+gender:newPTCA)
print(model.tables(lmc, "means"), digits = 3)
gen<-factor(gender)
newPT<-factor(newPTCA)
interaction.plot(gen,newPT,maxhr, type = "b", col = c(1,2)
                 , xlab = "gender", ylab = "mean maximum heart rate")

```
There is obvious interaction between the variables gender and newPTCA based new angioplasty or no new angioplasty.

d) Conduct the same analyses using regression. Compare and contrast the results of regression and ANOVA.
```{r}
summary(lm)
summary(lmc)
```
The results from regression are close to those from the anova.

e) Interpret R2 resulting from regression analyses within the context of the
study.

Only 0.5% of the variation within the response variable is explained by the model.

Part two. Using the stress echo data set posted earlier CCLE conduct an ANCOVA (Analysis of Covariances) and answer the following questions.

a) Using maximum heart rate as the outcome variable, base heart rate as the covariate, and ecg as the predictor, conduct ANCOVA. Report both outputs.

```{r}
ECG <- factor(ecg)
lm2 <- lm(maxhr~ECG+bhr)
summary(lm2)
aov(lm2)
```

b) Compute the coefficient of correlation between maximum heart rate and base heart rate and the scatter plot by group (normal ecg and equivocal ecg ??? equivocal meaning dubious or questionable); this way we can make sure that the pattern of relationship between the covariate (bhr) and outcome (maxheartrate) is similar for both groups.

```{r}
cor(maxhr,bhr)
par(mfrow=c(1,2))
plot(maxhr ~ ECG + bhr)
```

c) Interpret the coefficients associated with ecg and ecg:bhr. Determine whether bhr is a significant covariate or not.
Hints:

```{r}
lm2c <- lm(maxhr ~ ECG + bhr + ECG:bhr)
summary(lm2c)
```
bhr is a significant covariate base on the statistics.

Compute the mean of maxhr by ecg for a more accurate interpretation of the coefficient associated with ecg. You could also run SLR as a function of ecg to see this.
Part Three: Question 3 of Chapter 5

a)
The coefficient of the interaction term in model is statistically significant, becuase the p-value is less than 0.025 so that we have a significant interaction term. Also, the r-squared value increased by 5% for including the interaction term. 

b)
  Quality=b0 +b1EndofHarvest+b2Rain + b3 End of Harvest ?? Rain 
  Quality=5.16122 - 0.03145*EndofHarvest + 1.78670*Rain - 0.08314*End of Harvest ?? Rain
  (i) No unwanted rain at harvest (rain=0)
  -1 = 5.16122 - 0.03145*EndofHarvest
```{r}
-1/-0.03145
```

   (ii) Some unwanted rain at harvest(rain=1)
   
   -1=5.16122 - 0.03145*EndofHarvest + 1.78670 - 0.08314*End of Harvest
```{r}
-1/(-0.03145-0.08314)
```

Part Four: QUESTION 5 of Chapter 6
a) 



b)

```{r}
pgatour <- read.csv("~/Downloads/pgatour2006.csv", stringsAsFactors=FALSE)
attach(pgatour)
lm4 <- lm(PrizeMoney ~ DrivingAccuracy+GIR+PuttingAverage+BirdieConversion+
            SandSaves+Scrambling+PuttsPerRound)
summary(lm4)
par(mfrow=c(2,2))
plot(lm4)
```
There is a pattern in the residual vs fitted plot, which indicates the violation of random normal errors assumptions. The  Violations of the basic assumptions may invalidate our model. As a result, we choose to transform the model.
```{r}
lm5 <- lm(log(PrizeMoney) ~
            log(DrivingAccuracy)+log(GIR)+log(PuttingAverage)+log(BirdieConversion)+log(SandSaves)+log(Scrambling)+log(PuttsPerRound))
par(mfrow=c(2,2))
plot(lm5)
```

The second model has liitle or no violation of the basic assunptions. Thus, we take the transformed model instead of the original one.

c) 
```{r}
outlier <- which(cooks.distance(lm5)> 4/(nrow(pgatour)-2))
outlier
```

These points need further examination, because they are outliers and may have significant effect on the model. 

d) 

```{r}
library(car)
vif(lm5)
mmps(lm5)
```

Some of the variables have a vif value over 5 and some of the predictors are not quite inhancing the model. Thus, we should consider a simpler model compare to the full model.
e)

Because there may be multicorlinearity between the predictors, we cannot delete the predictors with insignificant t-values at the same time. We should start omitting predictors one by one instead of taking out the indignificant variables at the same time.