---
title: "101AHW8"
author: "JING LI"
date: "11/30/2016"
output: pdf_document
---

Credit data set:
Question 1:
Homework 8
??? Please download credit data from ccle week 10.

```{r}
credit <- read.csv("~/Downloads/Credit.csv", stringsAsFactors=FALSE)
credit$Balance <- credit$Balance+0.001
dim(credit)
head(credit)
attach(credit)
```

??? This data has 400 observations and 11 variables.
??? Use Balance as your response variable.
??? Split your data into 70% tranning and 30% testing using set.seed(654321)

```{r}
smp_size <- floor(0.70 * nrow(credit)) 
smp_size
set.seed(654321)
train_ind <- sample(seq_len(nrow(credit)), size = smp_size)
train <- credit[train_ind, ]
test <-  credit[-train_ind, ]
write.table(train, "~/Downloads/birthsTrain.txt", sep="\t")
write.table(test, "~/Downloads/birthsTest.txt", sep="\t")
head(train)
```

??? Create a full model with all the 10 predictors.

```{r}
creditm1<-lm(credit$Balance~.,data=credit[,2:11]) 
```

??? create a summary for your model and the anova for the full model

```{r}
summary(creditm1)
anova(creditm1)
```

??? Which predictors are significant? Report your R-Square.

Significant predictors: Income, Limit, Rating, Cards, and StudentsYes. The R-square from the full model is 0.9551.

??? Check diagnostics for your full model. Comment on violations if any. (don't forget the vif)

```{r}
par(mfrow=c(2,2)) 
plot(creditm1) 
library(alr3)
library(car)
vif(creditm1)
```

There are violations of the constant variances and normality. Vif values indicate that variables Limit and Rating have problems with colinearity, so we can drop one of them first and check the model again. 

```{r}
creditm2 <- lm(Balance ~ Income+Limit+Cards+Age+Education+Gender+Student+Married+Ethnicity, data = train)
summary(creditm2)
vif(creditm2)
par(mfrow=c(2,2)) 
plot(creditm2) 
```

??? Use inverse response plot to find the best transformation of the response variable.

```{r}
par(mfrow=c(1,1))
inverseResponsePlot(creditm2)

```


??? create mmps for your response variable and all your predictors.

```{r}
detach(credit)
attach(train)
```

??? Use powerTransform to find the best transformation(s) to be done on your variables.

```{r}
summary(powerTransform(cbind(train$Balance, Income, Limit, Cards, Age, Education)~1), data = train)
```

??? Conduct a new full model with the transformed variables. Check assumptions.

```{r}
BalanceT <- train$Balance^(1/3)
IncomeT <- train$Income^(1/5)
LimitT <- sqrt(Limit)
CardsT <- sqrt(Cards)
creditm3 <- lm(BalanceT ~ IncomeT +LimitT+CardsT+Age+Education+Gender+Student+Married+Ethnicity, data = train)
par(mfrow=c(2,2)) 
plot(creditm3) 
```


??? Use the step function with the backword method and AIC criteria to find the best subset of your predictors. Report your model and its summary using the best preditors only.

```{r}
backAIC <- step(creditm3,direction="backward", data=train)
creditm4 <- lm(BalanceT ~ IncomeT + LimitT + CardsT + Student + Ethnicity, data = train)
summary(creditm4)
```

??? Check assumptions.

```{r}
plot(creditm4)
```

??? Now use your testing data to create a SLR model between the (transformed predicted response) and the transformed response from the testing data.

```{r}
detach(train)
attach(test)
BalanceT <- train$Balance^(1/3)
IncomeT <- train$Income^(1/5)
LimitT <- sqrt(Limit)
CardsT <- sqrt(Cards)
Student_Yes <- rep(NA, 120)
for (i in 1:120){
  if(test$Student[i]=="Yes") Student_Yes[i] <- 1
  else  Student_Yes[i] <- 0
}
Ethnicity_Asian <- rep(NA, 120)
Ethnicity_Caucasian <- rep(NA, 120)
for (i in 1:120){
  if(test$Ethnicity[i]=="Asian") Ethnicity_Asian[i] <- 1
  else  Ethnicity_Asian[i] <- 0
}
for (i in 1:120){
  if(test$Ethnicity[i]=="Caucasian") Ethnicity_Caucasian[i] <- 1
  else  Ethnicity_Caucasian[i] <- 0
}
ypredtest<- 0.54793 +  -6.69626*test$Income +  0.27817*test$Limit + 0.31851*test$Cards + 3.21258*Student_Yes + 0.08469*Ethnicity_Asian + 0.49966*Ethnicity_Caucasian
creditm5 <- lm(ypredtest~Balance, data = test)
summary(creditm5)
plot(creditm5)
```

??? Create a scatter plot for your SLR and report its R-Square.

```{r}
par(mfrow=c(1,1)) 
plot(test$Balance,ypredtest)
detach(test)
```


The R-Square from the SLR is 0.874

??? Comment on your findings.

The predicted value from our training data fit quite well with the observations from the testing data. Overall, we have a satisfied model from our training data.




Part one. Refer to table one and answer the following questions.
What are the major areas in which masculine idealists and masculine moderates are significantly different from each other? Explain the results within context and mention which group is higher and which group is lower?

Self-reported health, College degree or more (percent), Prostate exam in past 12 months (percent), Flu shot in past 12 months (percent), Compliant with all three services (percent) are major areas in which masculine idealists and masculine moderateds are significantly different from each other. Masculinity idealists were less likely than moderates to have obtained preventive health care in the last 12 months. Masculine idealists have lower percentages in these areas.


Part two. Refer to table two and answer the following questions
??? Interpret the odds of 1.65 associated with college degree
If College degree or more (percent) increases one percent, the odds of the person being Compliant with All Three Services increases by 1.65.

??? Interpret the confidence interval of (1.20, 2.27) associated with college degree.

We are 95% confident that for one unit of increase in X or College degree or more (percent) the odds of the person being Compliant with All Three Services increases from 1.20 to 2.27. 

??? Interpret the odds of 1.02 associated with household income

Each $10,000 increase in household income is associated with 2 percent increase in the likelihood of obtaining a Physical Examination.
