---
title: "102AHW1"
author: "JING LI"
date: "September 29, 2016"
output: html_document
---

---
title: "Stats 102A (Computational Statistics) - Homework 1 - Data Structures"
author: "Put Your Name Here"
date: 
output: html_document
---

_This is your first homework assignment. It is an R Markdown document. You will modify this file with your answers and responses. Please preserve all text that is italicized. This will signal to the course reader that the italicized text is part of the homework prompt and is not to be graded. Everything else can (probably should) be modified to answer the question. If you must emphasize text in your answers, please use boldface, rather than italics._

_This homework assignment is a bit more conceptual than future assignments. It does, however, highlight some important things to understand about how R stores and handles values._

1. _Working with vectors_
    a. _Create a vector of the positive odd integers less than 40_
    b. _remove the values greater than 20 and less than 30_
    c. _Find the variance of the remaining set of values_

```{r}
#1a
seq(1, 40, by = 2)   # 1a been done to get you started
#1b
c(seq(1,20, by=2), seq(31,40, by=2))
#1c
var(c(seq(1,20, by=2), seq(31,40, by=2)))
```


2. _Coercion: For each of the following, explain what type of output you will receive and why R is producing that output._
    a. `c(0, TRUE)`
    b. `c("f", 1)`
    c. `c(list(1), "b")`
    d. `c(FALSE, 1L)`
```{r}
#2a
typeof(c(0, TRUE))
#2b
typeof(c("f", 1))
#2c
typeof(c(list(1), "b"))
#2d
typeof(c(FALSE, 1L))
```


3. _What is the difference between NULL, NA, and NaN?_

NULL: NULL object NA: a missing value NaN:Not a Number

4. _What is the difference between logical(0) and NULL?_

```{r}
typeof(logical(0))
typeof(NULL)
```


5. _A vector `c(TRUE, FALSE)` is a logical vector. Other than `TRUE` or `FALSE`, what can you 'put' into the vector so that it remains a logical vector and does not get coerced into another class?_

```{r}
typeof(c(TRUE, FALSE, NA, NULL))
```

We can put NA and NULL into the vector, and the vector will still has the type logical.

6. _Why do you need to use `unlist()` to convert a list to an atomic vector? Why doesn't `as.vector()` work?_

Because list is a generic vector, as.vector will not alter a generic vector to a atomic vector. However, unlist will coerce the type of the vector according to the hierachy.

7. _What will `c(4:7) * c(2:4)` produce? Briefly, why?_

The output will be: 8 15 24 14 (4*2 5*3 6*4 7*2)

8. _What happens to a factor when you modify its levels?_

```{r}
f1 <- factor(LETTERS)
levels(f1) <- rev(levels(f1))
```

Both the levels and the contents are reversed in this case.

9. _What does this code do? How do f2 and f3 differ from f1?_

```{r}
f2 <- rev(factor(LETTERS))
f2
f3 <- factor(LETTERS, levels = rev(LETTERS))
f3
```

f2 has the reversed letters as contents and the ordered letters as levels, while f3 has the reversed letters as levels and the ordered letters as contents.

10. _I have created matrix AA, a 10 x 10 matrix with the numbers 1 through 100. Sum all the even rows of the second column of the matrix. Do this in one line of code._

```{r}
AA <- matrix(1:100, nrow=10)
# write your answer here
sum(AA[seq(2,10, by=2), 2])
```

11. _What attributes does a data frame possess?_

class, comment, dim, dimnames, names, row.names and tsp

12. _What does as.matrix() do when applied to a data frame with columns of different types?_

```{r}
df1 <- data.frame(col1=c(1,2,3), col2=c('a','b','c'))
df1
as.matrix(df1)
```
The content of the data frame will be coerced into the least restrictive objects, in this case: characters.
13. _Can you have a data frame with 0 rows? What about 0 columns?_

```{r}
df1 <- data.frame(col1=integer(), col2=double())
df1
df2 <- data.frame(row.names = c('r1','r2','r3'))
df2
df3 <- data.frame()
df3
```
YES

14. _What is the difference between the assignment operators `<-` and `<<-`? Look up help(assignOps)_

The operators <- and = assign into the environment in which they are evaluated, while the operators <<- and ->> are normally only used in functions, and cause a search to made through parent environments for an existing definition of the variable being assigned. If such a variable is found (and its binding is not locked) then its value is redefined, otherwise assignment takes place in the global environment. 

15. _Some R Vocabulary_

To be proficient in R Programming, you need to have a good working vocabulary. You should familiarize yourself with all of the following functions. You don't have to write anything or 'do' anything for this exercise, other than do the appropriate reading to understand the usage of these functions.

Please take some time to look up the help files for each of the following functions or operators. I strongly encourage you to read through and try out some of the examples at the bottom of the help pages to understand their usage. This activity should take you a couple hours to complete. Credit to Hadley Wickham for compiling an excellent list, from which I have drawn a few functions. His complete list can be found here:_ <http://adv-r.had.co.nz/Vocabulary.html>

```{r, eval=FALSE}
str

match, %in%
subset
with

# comparison
all.equal, identical
==, !=, >, >=, <, <= # help(comparison)
is.na
any


# basic math
*, +, -, /, ^, %%, %/% # help(arithmetic)
abs, sign
ceiling, floor, round, trunc, signif # help(Round)
max, min, prod, sum
cummax, cummin, cumprod, cumsum
diff

# Logical & sets 
&, |, !, xor # help(Logic)
all, any
intersect, union, setdiff, setequal
which

# Vectors and matrices
c, matrix
# automatic coercion rules character > numeric > logical
length, dim, ncol, nrow
cbind, rbind
names, colnames, rownames
t
diag
as.matrix, data.matrix

# Making vectors 
c
rep
seq
rev
sample
(is/as).(character/numeric/logical/...)

# Lists & data.frames 
list, unlist
data.frame

# Factors 
factor, levels, nlevels
reorder, relevel

# Array manipulation
array
dim
dimnames

# Ordering
merge
order
rank
sort

# inputs / outputs
cat
print
paste
readline
```